Building Markov State Model for a Periodically Driven Non-Equilibrium System

### 2014-09-06 writedown/period.noneq.msm/paper.tex
backup

\section{Introduction}
Non-equilibrium, especically periodically driven system. Interesting.

MSM tools for analyzing, provide profund understanding.

@@

$T$-periodic external field $E(t)$. $\beta$ is the inversed temperature,

@@

  \label{eq:disc-2}

@@

Following Ref.~\cite{latorre2011structure}, the Fokker-Planck Eq.~\ref{eq:disc-2}

@@

\recheck{Moreover,

@@

One possible approximation to $L(t)$ is via the following forward finite difference scheme:
  \label{eqn:tmp4}
  L_{ij}(t) \approx \frac{1}{\tau}
  \,[\, \prob (\vect X_{t+\tau} = i \vert \vect X_{t} = j) - \delta_{ij} \,]
where $\vect X_t$ denotes the Markov jump process generated by $\vect L(t)$, 
and $\tau$ is the disretization time-step.
In analogy to \eqref{eq:disc-2}, the Markov jump process generated by

@@

  \label{eq:disc-7}

@@

  \label{eq:disc-9}

@@

@@

of $\ml(t)$, in particular $p_i(t)=\prob[x_t\in \Omega_i]$.

@@

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. $\vect Y$ satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L(t+T)^T \vect Y(t)=\vect L(t)^T\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T),
\end{equation}
and in particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T).
\end{align}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T]$.

Since $\myphi(T)$ is a stochastic matrix, the spectrum $\sigma(\myphi(T))$
of the matrix $\myphi(T)$ is contained in the circle in
the complex plane, i.e., each eigenvalue $\lambda\in \sigma(\myphi(T))$
satisfies $|\lambda|\le 1$. Furthermore $1\in\sigma(\myphi(T))$ is an
eigenvalue with left eigenvector $\vect e$ and a right eigenvector $\mymu$
satisfying
$\myphi(T)\mymu=\mymu$.
From now on, we assume $\myphi(T)$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). In this case $\mymu$ is the stationary measure in the sense that
\begin{align}
\label{eq:floq-14}  
\myphi(mT) \mymu = \mymu,\quad m=0,1,2,\ldots,
\end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)=\vect p_0$ by the process satisfies
\begin{align}
\label{eq:floq-15}  
\myphi(mT)\vect p_0\to \mymu,\quad m\to\infty.
\end{align}

@@

\section{Numerical example:
  Alanine dipeptide under oscillatory electric field}

If the spacial discretization is fine enough, the Markov jump
process~\eqref{eq:disc-7} is a good approximation to the original
MD~\eqref{eq:disc-1}.  In practice, it is firstly difficult to predict
how many discret set is fine enough; Secondly, since the
total configurational dimension of the system is $3N$ ($N$ being the
number of atoms), it is prohibitive to do a very fine discretization
over all DOFs for most of systems of practical interest.  In the
simulation of biomolecules, in most cases it is reasonable to assume a
time-scale separation in the system.
For example, the conformational
change of the molecules is much faster than the covalent bond
vibrations. 
% It should be noted that we do NOT assume that the fast time-scales are decoupled with the slow time-scales~\footnote{If the fast time-scales
%   are decoupled from the slow ones, then the story would become much simpler,
% however, this situation is lack of generality, we will not discuss.}.
In this context,
it is usually possible to pick a relatively small number of discrete sets that
correctly describe the slow dynamics in the sense that
the original dynamics fully relaxes in every set
within the discretization time-step (denoted by $\tau$). Or equivalently the discretized
dynamics~\eqref{eq:disc-7} is a good approximation to~\eqref{eq:disc-1}.
One possible way to define the discretized set is to firstly find few collective
variables, and then to discretize the dynamics as fine as possible only on these
variables either by uniform or adaptive
discretization~\cite{chodera2007automatic, prinz2011markov}.
This process also suggests that the choice of the lag-time should be 
shorter than the dominant implied timescale so that they can be resolved,
and be longer than the fast time-scales to relax the unresolved dynamics
within the discretized sets.
However, it is in general difficult to tell in priori how to choose the collect
variables, how fine the discretization should be and how large
the time-steps should be.
For large system, these questions are usually
not trivial to answer.

@@

To illustrate how the procedure works in practice, it is therefore
useful to investigate the discussed approximations by a numerical
example.  We take the alanine dipeptide system under an oscillatory EF,
as an example, the non-equilibrium MD simulation of which was
extensively studied by Ref.~\cite{wang2014exploring}.  
The alanine dipeptide was put into the local thermostating
environment, and was driven by a periodic electric field with period
$T = 10$~ps. The 20,000 branching trajectories started from 20,000
initial configurations that samples the equilibrium distribution.
\redc{Write a lot of details on non-equilibrium MD.}  The branching
trajectories were each 4000~ps long, and the system reaches
non-equilibrium steady state in roughly 300~ps.

We choose the
two dihedral angles $\phi$ and $\psi$ as collective variables (see
Fig.~\ref{fig:tmp1}), and the discretization is a uniform division of
the $\phi$--$\psi$ plane. We denote the number of division on each
dihedral by $K$. In the following subsections, we investigate the discretization
quality firstly by considering the choice of $K$ and time-step $\tau$. Then
the discretized dynamics is compared with original (non-equilibrium brute force simulation)
via considering the steady state probability density, the first mean hitting time
and the forward and backward committors.

\subsection{Discretizing the dihedral angles}
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t010/discrete/fig-cg-prob.eps}  
  \caption{The time-dependent probability $\mathbb
    P\big(\phi\in[0,180), \psi\in [0,180)\big)$.  The brute force
    non-equilibrium MD simulation is compared with different
    discretization methods. The red shadow region indicates the
    statsitical uncertainty of the brute force simulation.}
  \label{fig:tmp2}
\end{figure}

We estimate the discretized generator $L(t), \ t\in[0,T)$
from non-equilibrium trajectories generated by $\ml(t)$ in time interval $[t_1,
t_2]$.
Due to the periodicity of $\ml(t)$, if the discretized dynamics approximate the original
dynamics good enough, generator  $L(t)$ should not depend on in which interval
it is estimated, provided that the initial state of the system
is not very far from the steady state.
This is a good indicator of the discretization quality.
We compute $L(t)$ by two discretizations $K=2$ and $K=20$, and two
choices of time intervals $[0, 80]$~ps and $[320, 400]$~ps, and then
compare the time-dependent
probability $\prob\big(\phi_t\in[0,180), \psi_t\in [0,180)\big)$
with the brute force result in
Fig.~\ref{fig:tmp2}. 
Using $K=2$ the dynamics depends on the time interval used for
calculating the generator: using time interval $[0, 80]$~ps the discretized
dynamics deviates the brute force result,
while using time interval $[320, 400]$~ps the discretized dynamics can only
reproduces the brute force result after 300~ps.  This therefore indicates poor 
approximations to the original dynamics with $K=2$. The reason is that the
discretization is too coarse so that the dynamics cannot be fully
equilibriated within the lag-time $\tau$ in each discretized set,
therefore, the discretization presents state dependency.  For
$K=20$, the discretized dynamics does not depend on the time interval of
calculating the generator, and is consistent with the brute force
non-equilibrium simulation within the error bar. Therefore, in this paper we use $K=20$
to discretize  the dihedral angle space of the alanine dipeptide.
For a good statistical accuracy, if not stated otherwise,
we will use the full trajectories, i.e.~a time interval of
$[0,4000]$~ps for estimating the discretized dynamics.

\subsection{The choice of lag-time $\tau$}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t010/discrete/fig-cg-prob-tau.eps}  
  \caption{The time-dependent probability $\mathbb
    P\big(\phi\in[0,180), \psi\in [0,180)\big)$.  The brute force
    non-equilibrium MD simulation is compared with different
    discretization methods. The red shadow region indicates the
    statsitical uncertainty of the brute force simulation.}
  \label{fig:tmp3}
\end{figure}

As discussed before, the lag-time should be chosen a value that lies
in the spectrum gap of the original dynamics, so that it resolves the
slow dynamics, and at the same time relaxes the fast dynamics. In
practice, it is very difficult to estimate the lag-time in
priori. Therefore, we consider 
different choices of $\tau$ (0.5, 1.0, 2.0 and 5.0~ps)
discretizing the same original dynamics, and compare the
time-dependent probability $\prob\big(\phi\in[0,180), \psi\in
[0,180)\big)$ calculated from different time discretization with the brute force simulation
of the original dynamics
(see Fig.~\ref{fig:tmp3}).  All cases use idential dihedral angle discretization: $K=20$.
It is clear that when lag-time is close to the period (10~ps), the
discretized dynamics cannot resolve the probability change within a
period. However, it is surprising  that the large lag-times are able to capture the
the overall log time behavior of the probability.
We observe no significant difference between $\tau=0.5$ and
$\tau=1.0$~ps, which means the discretized dynamics is not very sensitive
to the choice of $\tau$. 
We do not investigate the small $\tau$ limit, because saving trajectories
more frequently requires more hard disk space, and is usually not preferred.

@@

\begin{figure}
  \centering  
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.40.steadyDist//fig-dist.eps}
  % \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-dist-msm.eps}
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.40/fig-floquet-vec-1.eps}
  \caption{The steady state distribution calculated by brute force non-equilibrium is compared with that by the discretized dynamics.}
  \label{fig:num-1}
\end{figure}

An important check of the discretized dynamics is that if it reporduces the steady probability
density $\mymu$ of $\myphi(T)$.
To make it comparible to
the free energy in equilibrium case, we take the logrithm of the properties, i.e.
\begin{align}
  \label{eq:num-tmp1}
  F_{\textrm{st}}(\phi,\psi) = \lim_{m\rightarrow\infty} -k_B\mathcal T \log p(\phi,\psi,mT),
\end{align}
where $k_B$ is the Boltzmann constant and $\mathcal T$ is the temperature of the system.
The result is given in Fig.~\ref{fig:num-1}. A good consistency between the brute force
MD simulation and the discretized dynamics is observed.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t040/cluster.marco/fig-cluster-2.eps}
  \caption{The coreset identification. Different colors indicate different coresets: $C_{\confaa}$ (yellow), $C_\beta$ (red) and $C_{\confc}$ (green).
    The blue color means out of any coreset. (By Marco)}
  \label{fig:cluster}
\end{figure}

\redc{Write how to detact the core sets for irreversible Markovian process}.

@@

\subsection{First mean hitting time}

Since the largest first mean hitting time (starting fom coreset $\confaa$ to $\confc$) is
longer than 600~ps, if we use the trajectories of length 4000~ps, the
results will be biased. Therefore, we use 100 non-equilibrium trajectories
of $2\times 10^5$~ps instead.
The first mean hitting time is presented in Fig.~\ref{fig:num-6}.

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-3.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-msm-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-msm-3.eps}
  \caption{The first mean htting time comparsion between the brute
    force simulation (first row) and the discretized dynamics (second row).  From left
    to right are first mean hitting time to coresets $C_{\confc}$,
    $C_{\confaa}$ and $C_{\beta}$, respectively}
  \label{fig:num-6}
\end{figure}

\subsection{Forward and backward committors}

The forward committor $q^\fwd_i(\phi,\psi)$ of a coreset $C_i,\
i\in\{\confaa, \beta, \confc\}$ is defined as the probability of
visiting coreset $C_i$ next conditioned on being at configuration
$(\phi,\psi)$.  The backward committor $q^\bwd_i(\phi,\psi)$ of a
coreset $C_i,\ i\in\{\confaa, \beta, \confc\}$ is defined as the
probability of last coming from $C_i$ conditioned on being at
configuration $(\phi,\psi)$.  For reversible Markov processes, the
forward and backward committors are identical, however, it is in
general not the case for irreversible processes. Please see
Fig.~\ref{fig:num-3}--\ref{fig:num-5} for the committors.

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-1.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-msm-1.eps}
  \caption{The forward $q^\fwd_{\confc}$ and backward committors
    $q^\bwd_{\confc}$ computed by discretized dynamics (second row) is compared with
    those computed by brute force non-equilibrium simulations (first
    row).}
  \label{fig:num-3}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-2.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-msm-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-msm-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-msm-2.eps}
  \caption{The forward $q^\fwd_{\confaa}$ and backward
    $q^\bwd_{\confaa}$ committors computed by discretized dynamics (second row) is
    compared with those computed by brute force non-equilibrium
    simulations (first row).}
  \label{fig:num-4}
\end{figure}

>>>>

\section{Introduction}
Non-equilibrium, especially periodically driven system. Interesting.

MSM tools for analyzing, provide profound understanding.

@@

$T$-periodic external field $E(t)$. $\beta$ is the inverse temperature,

@@

  \label{eq:disc-fp}

@@

Following Ref.~\cite{latorre2011structure}, the Fokker-Planck Eq.~\ref{eq:disc-fp}

@@

{Moreover,

@@

In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by

@@

  \label{eq:disc-master}

@@

  \label{eq:disc-master-phi}

@@

where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

of $\ml(t)$, in particular $p_i(t)=\prob(x_t\in \Omega_i)$.

@@

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. $\vect Y$ satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L(t+T)^T \vect Y(t)=\vect L(t)^T\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T]$.  
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
and for the probability at integral period we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0).
\end{align}

Since $\myphi(T)$ is a stochastic matrix, the spectrum $\sigma(\myphi(T))$
of the matrix $\myphi(T)$ is contained in the circle in
the complex plane, i.e., each eigenvalue $\lambda\in \sigma(\myphi(T))$
satisfies $|\lambda|\le 1$. Furthermore $1\in\sigma(\myphi(T))$ is an
eigenvalue with left eigenvector $\vect e$ and a right eigenvector $\mymu$
satisfying
$\myphi(T)\mymu=\mymu$.
From now on, we assume $\myphi(T)$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). In this case $\mymu$ is the stationary measure in the sense that
\begin{align}
\label{eq:floq-14}  
\myphi(mT) \mymu = \mymu,\quad m=0,1,2,\ldots,
\end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
\begin{align}
\label{eq:floq-15}  
\myphi(mT)\vect p(0)\to \mymu,\quad m\to\infty.
\end{align}

Using the Floquet theorem, the time-inhomogeneous Markov process $X_t$
is simplified into a \emph{time-homogeneous} (not necessarily
reversible) Markov jump process $\tilde X_{m} = X_{mT}, \ m\in\mathbb
N$.
Therefore, comparing with $X_t$, we prefer to consider the process $\tilde X_{m}$
that helps understand the original dynamics $x_t$, because the powerful theories and
computational tools for time-homogeneous processes can be directly applied.
There is no doubt that information within one period is lost by using this temporal
discretization, however, information regarding the long-term behaviors
of the system would be satisfactorily described as long as their time-scales
are much longer than the period.
At the same time,
the computational cost of generating $\tilde X_{m}$ is mush less demanding
than the brute force simulations of NEMD, which implies lower
statistically uncertainty in calculating the observables of interest.
% due to the discretization in both the spacial and temperal directions.
% In the numercal example in Sec.~\ref{sec:alanine}, we
% compute the

@@

\section{Numerical example:
  Alanine dipeptide under oscillatory electric field}
\label{sec:alanine}

If the spacial discretization is fine enough, the Markov jump
process~\eqref{eq:disc-master} is a good approximation to the original
MD~\eqref{eq:disc-1}.  In practice, it is difficult to predict
how many discrete set is fine enough; Moreover, since the
total configurational dimension of the system is $3N$ ($N$ being the
number of atoms), it is prohibitive to do a very fine discretization
over all DOFs for most of systems of practical interest.
In the systems that possess a
time-scale separation, it is possible, in most cases, to pick up
a characteristic time (or lag-time) $\tau$ that distinguishes the slow
and fast time-scales.
For example, the conformational
transitions of a large biomolecule is much slower than the covalent bond
vibrations. 
% It should be noted that we do NOT assume that the fast time-scales are decoupled with the slow time-scales~\footnote{If the fast time-scales
%   are decoupled from the slow ones, then the story would become much simpler,
% however, this situation is lack of generality, we will not discuss.}.
In this context,
it is usually possible to construct a
relatively small number of discrete sets that
correctly describe the slow  dynamics, and 
in each set the fast dynamics relaxes within the lag-time $\tau$.
Then if the discretized dynamics~\eqref{eq:disc-master}
reproduces the slow time-scales and the corresponding transitions
of the original dynamics~\eqref{eq:disc-1},
the former is considered to be a good approximation of the later.
% Then the 
% dynamics~\eqref{eq:disc-master} built on top of the discretization
% is a good approximation to~\eqref{eq:disc-1}
% in the sense of reproducing the slow dynamics.
One possible way to define the discretized set is firstly to find a few collective
variables, and then to discretize the dynamics as fine as possible only on these
variables either by uniform or adaptive
discretization~\cite{chodera2007automatic, prinz2011markov}.
% This process also suggests that the choice of the lag-time should be 
% shorter than the dominant implied timescale so that they can be resolved,
% and be longer than the fast time-scales to relax the unresolved dynamics
% within the discretized sets.
However, it is difficult to give a general answer in prior regarding, e.g.
how large is the lag-time,
how to choose the collect variables, how fine the discretization should be.
For large or high dimensional systems, these questions usually become non-trivial.

@@

To illustrate how the discretization works in practice, it is therefore
useful to investigate the discussed approximations by a numerical
example.  We take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied by Ref.~\cite{wang2014exploring}.  
The alanine dipeptide was put into the local thermostating
environment, and was driven by a periodic electric field with period
$T = 10$~ps. The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
\redc{Write a lot of details on NEMD.}  The branching
trajectories were each 4000~ps long, and the system reaches
non-equilibrium stationary state in roughly 300~ps.

We choose the
two dihedral angles $\phi$ and $\psi$ as collective variables (see
Fig.~\ref{fig:tmp1}), and the discretization is a uniform division of
the $\phi$--$\psi$ plane. We denote the number of division on each
dihedral by $K$, then the discretized sets are
$\{\Omega_i\},\ i\in S = \{1,\cdots,K^2\}$.
One possible approximation to
the discretized generator $L(t)$ is via
the following forward finite difference scheme:
\begin{align}
  \label{eqn:tmp4}
  L_{ij}(t) \approx \frac{1}{\tau}
  \,[\, \prob (\vect X_{t+\tau} = i \vert \vect X_{t} = j) - \delta_{ij} \,],
  \quad i,j\in S
\end{align}
and $\tau$ is the lag-time.
In the following subsections~\ref{sec:alanine-disc} and \ref{sec:alanine-tau},
we investigate the discretization
quality with respect to  the choice of $K$ and  lag-time $\tau$.
Then
the discretized dynamics generated by the Floquet transition matrix
is compared with original NEMD simulation
via considering the stationary probability density, the first mean hitting time
and the forward and backward committors.

\subsection{Discretizing the dihedral angles}
\label{sec:alanine-disc}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t010/discrete/fig-cg-prob.eps}  
  \caption{The time-dependent probability $\mathbb
    P\big(\phi\in[0,180), \psi\in [0,180)\big)$.  The NEMD simulation is compared with different
    discretization methods. The red shadow region indicates the
    statistical uncertainty of the NEMD simulation.}
  \label{fig:tmp2}
\end{figure}

We estimate the discretized generator $L(t), \ t\in[0,T)$
from NEMD trajectories generated by $\ml(t)$ in time interval $[t_1,
t_2]$.
Due to the periodicity of $\ml(t)$, if the discretized dynamics approximate the original
dynamics good enough, generator  $L(t)$ should not depend on in which interval
it is estimated, provided that the initial state of the system
is not very far from the stationary state at long-time limit.
Therefore, this is an indicator for calibrating the discretization quality.
We compute $L(t)$ by two discretizations $K=2$ and $K=20$, and two
choices of time intervals $[0, 80]$~ps and $[320, 400]$~ps, and then
compare the time-dependent
probability $\prob\big(\phi_t\in[0,180), \psi_t\in [0,180)\big)$
with the NEMD result in
Fig.~\ref{fig:tmp2}.
The time derivative of the master equation~\eqref{eq:disc-master}
is discretized by the forward Euler scheme with a time step of 0.5~ps.
Using $K=2$ the dynamics depends on the time interval used for
calculating the generator: using time interval $[0, 80]$~ps the discretized
dynamics deviates from the NEMD result,
while using time interval $[320, 400]$~ps the discretized dynamics can only
reproduces the NEMD result after 300~ps.  This therefore indicates poor 
approximations to the original dynamics with $K=2$. The reason is that the
discretization is too coarse so that the dynamics cannot be fully
equilibriated within the lag-time $\tau$ in each discretized set,
therefore, the discretization presents state dependency.  For
$K=20$, the discretized dynamics does not depend on the time interval of
calculating the generator, and is consistent with the
NEMD simulation within the error bar. Therefore, throughout this paper we use $K=20$
to discretize  the dihedral angle space of the alanine dipeptide.
For a good statistical accuracy, if not stated otherwise,
we will use the full trajectories, i.e.~a time interval of
$[0,4000]$~ps for estimating the discretized generator $L(t)$.

\subsection{The choice of lag-time $\tau$}
\label{sec:alanine-tau}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t010/discrete/fig-cg-prob-tau.eps}  
  \caption{The time-dependent probability $\mathbb
    P\big(\phi\in[0,180), \psi\in [0,180)\big)$.  The NEMD simulation is compared with different
    discretization methods. The red shadow region indicates the
    statistical uncertainty of the NEMD simulation.}
  \label{fig:tmp3}
\end{figure}

As discussed before, the lag-time should be chosen a value that lies
in the spectrum gap of the original dynamics, so that it resolves the
slow dynamics, and at the same time relaxes the fast dynamics. In
practice, it is very difficult to estimate the lag-time in
prior. Therefore, we consider 
different choices of $\tau$ (0.5, 1.0, 2.0 and 5.0~ps)
discretizing the same original dynamics, and compare the
time-dependent probability $\prob\big(\phi\in[0,180), \psi\in
[0,180)\big)$ calculated from different time discretization with the NEMD simulation
of the original dynamics
(see Fig.~\ref{fig:tmp3}).  All cases use identical dihedral angle discretization: $K=20$.
It is clear that when lag-time is close to the period (10~ps), the
discretized dynamics cannot resolve the probability change within a
period. However, it is surprising  that the large lag-times are able to capture the
the overall long time behavior of the original dynamics.
We observe no significant difference between $\tau=0.5$ and
$\tau=1.0$~ps, which means the discretized dynamics is not very sensitive
to the choice of $\tau$.
Therefore, throughout this paper $\tau=0.5$~ps will be used.
We do not investigate the small $\tau$ limit, because saving trajectories
more frequently requires more hard disk space, and is usually not preferred in practice.

@@

\begin{figure}
  \centering  
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3.steadyDist//fig-dist.eps}
  % \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-dist-msm.eps}
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3/fig-floquet-vec-1.eps}
  \caption{The color scale plot of the logarithm stationary probability
    of (a) the NEMD  and (b)
    the Floquet matrix $\Phi(T)$.
  }
  \label{fig:num-1}
\end{figure}

Comparing with the time-homogeneous Markov processes, there are not as
many tools for understanding the time-inhomogeneous Markov process.
Therefore, in the following subsections,
we consider the time-homogeneous process $\tilde X_{m}$ generated by
the Floquet transition matrix $P = \Phi^T(T)$, and investigate if it
reproduces the properties of the original  non-equilibrium process $x_t$.
In this context, only the configurations at the  integral periods $mT$ along the original process
are taken into consider.

An important check is the consistency between the
stationary probability density of $\Phi(T)$ (i.e.~the leading eigenvector $\mu$) and that 
estimated form the original NEMD simulation, along which only integral periods $mT$ are considered:
\begin{align}
  \label{eq:num-tmp1}
  \rho_{\textrm{st}}(\phi,\psi) = \lim_{m\rightarrow\infty} \rho (\phi,\psi,mT),
\end{align}
On each NEMD branching trajectory, the very beginning 320~ps is discarded, and
the rest of the trajectory in $[320,4000]$~ps is used to estimate
the stationary probability. $\Phi(T)$ is computed by solving~\eqref{eq:disc-master-phi}
in one period, where the discretized generator $L(t)$ was estimated
from the full NEMD trajectories of $4000$~ps long.
Moreover, to make it comparable to the free energy in equilibrium case, we take
the logarithm of the properties, i.e.~$F_{\textrm{st}}(\phi,\psi)=
-k_B\mathcal T\log \rho_{\textrm{st}}(\phi,\psi)$
for NEMD and $F_{\textrm{st}}(\phi,\psi)=
-k_B\mathcal T\log \mu(\phi,\psi)$ for $\Phi(T)$, where $k_B$ is the
Boltzmann constant and $\mathcal T$ is the temperature of the system.
% Under the aformentioned discretization, the stationary distribution is
% defined for the each bin of the dihedral angle space, i.e.~$F_{pq} =
% \int_{ph}^{(p+1)h}d\phi\int_{qh}^{(q+1)h}d\psi
% F_{\textrm{st}}(\phi,\psi), \ 0\leq p,q<K-1$, where $h = 360/K$ being
% the size of the bin, $p$ and $q$ here being the bin indexes.
The
results are  compared in Fig.~\ref{fig:num-1}. A good consistency between
the  NEMD simulation and 
$\Phi(T)$ is observed.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3/fig-cluster.eps}
  \caption{The coreset identification. Different colors indicate different coresets: $C_{\confaa}$ (green), $C_\beta$ (yellow) and $C_{\confc}$ (red).
    The blue color means out of any coreset. (By Marco)}
  \label{fig:cluster}
\end{figure}

\redc{Write how to detect the core sets for irreversible Markov process}.

@@

\subsection{First mean hitting time}
\label{sec:alanine-fmht}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-1.eps}\\
  \vskip -.5cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-2.eps}\\
  \vskip -.5cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-3.eps}\\
  \caption{The first mean hitting time (FMHT) comparison between the NEMD
    simulation (first column) and the discretized dynamics (second column).  From up
    to down are first mean hitting time to coresets $C_{\confaa}$ and $C_{\beta}$
    and $C_{\confc}$, respectively}
  \label{fig:num-6}
\end{figure}

The first mean hitting time, as a function of the dihedral angles $(\phi,\psi)$,
is defined by the
averaged first time needed for hitting a certain coreset
$i\in\{\alpha_R,\beta,\alpha_L\}$, conditioned on starting from the
conformation $(\phi,\psi)$.  Since the largest first mean hitting
time (starting from coreset $\confaa$ to $\confc$) is longer than
600~ps, if we use the NEMD trajectories of length 4000~ps,
the results will be biased. Therefore, we use 100 NEMD
trajectories of $2\times 10^5$~ps instead. We also compute the first mean hitting
time of 
a  $10^9$~ps long (being $T=10$~ps, $m = 1,\cdots, 10^8$)
discretized trajectory $\tilde
X_{m}$ that is generated by the Floquet transition matrix $P=\Phi^T(T)$.
The first mean hitting time is presented in Fig.~\ref{fig:num-6}.  The
good consistency between the NEMD and the discretized Markov process
$\tilde X_{m}$ indicates a good approximation quality.
Since the length of the discretized process is substantially longer than
the total length of NEMD
simulation ($10^9$~ps v.s.~$2\times10^7$~ps), the statistically uncertainty
is much smaller. 
However, one should notice that systematic errors
are introduced during the discretization (estimating $L(t)$), and 
its potential influence on the observables,
e.g.~the first mean hitting time, is not straightforward to estimate.
This topic is out of the scope of the current paper,
and will not be discussed in detail.
In summary, very long discretized trajectories helps in calculating the observables
in a smoother  (less statistical error) but not necessarily more accurate manner.
On the other hand, the computational cost of
discretized process, if the cost for estimating $\Phi(T)$ is not included, is
essentially cheaper than NEMD:
it took roughly 10 minutes on one core of an Intel Xeon E31245 CPU for generating the discretized trajectory, while the
NEMD trajectories took $1.6\times 10^4$ core hours for Intel Xeon E5-4650 CPUs.

\subsection{Forward and backward committors}
\label{sec:alanine-committor}
\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-1.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-1.eps}
  \caption{The forward $q^\fwd_{\confc}$ and backward committors
    $q^\bwd_{\confc}$ computed by discretized dynamics (second row) is compared with
    those computed by NEMD simulations (first
    row).}
  \label{fig:num-3}
\end{figure}
